Debugging by \textit{lastChange}

### 2010-11-01 trunk/papers/pldi2011/sigplanconf-template.tex
Claude's comments (first series).

According to \cite{LaToza}, developers spend about fifty percent of
their time debugging. To fix a bug, developers typically reproduce and
monitor the buggy execution several times to understand the program's
unexpected behavior. Trial and error, guess-work, and analyzing large
and complicated data make debugging difficult and time-consuming. Any
enhancement to debugging techniques and tools can save developers'
time, reducing development costs and improve programs' quality.

A common strategy for locating defects starts from bug symptoms and
works backwards, moving from a point in the program execution where a
value appears to be incorrect back to the point were that value was
set.  There are two conventional approaches used: breakpoint-based and
log-based debugging. Both approaches require tedious steps of
selecting data to be collected, collecting the data, then analyzing
the results.

In breakpoint debugging, developers select data to be collected by
searching through source files and setting breakpoints. To determine
where a value was set incorrectly, a developer must list and set
breakpoints at all possible points of the change for the value. At
every breakpoint the developer must determine if the location is in
fact related to the questionable value change then study the complex
debugger user interface and memorize values or manually collect
data. As the number of breakpoint hits increases, the process of
checking the program state, collecting data and resuming the execution
becomes cumbersome.

In log-based debugging, developers select data to be collected by
inserting statements for all points of possible change.  While in
breakpoint-based debugging, the whole program state is available to
developer, in log-based debugging, developer has to decide about data
should be collected when inserts the log statement. It is very common
that developer has to repeat this step several times due to
insufficient collected data, or to wait a long time because too much
data is recorded. Once the adequate data is collected, it still
requires analyzing and understanding. Developers usually end up in
dealing with long log files and analyzing huge collected data.  None
of these approaches assist developer in finding origins to a wrong
value.

To address these problems, we propose a new functionality in
debuggers, \textit{lastChange}, which locates the origin of a wrong
value. It uses program re-execution similar to the practice
in using breakpoints and log statements.

@@

We refer to the breakpoint used in \textit{lastChange} as the
\textit{reproduction point}: this is a point where the developer knows
reproduces the bug and it is the point were the \textit{lastChange}
algorithm ends. The ability to identify this reproduction point, i.e.
 a test case and a breakpoint after the bug is appears, is the only
prerequisite for \textit{lastChange} functionality.  This means
\textit{lastChange} can be implemented on top of a conventional
breakpoint debugger and it does not require a special environment to
create identical, instruction by instruction, re-executions.

Our contribution in this paper is the algorithm \textit{lastChange},
which locates the last place a value is changed, gathers other values
from that execution point, and allows \textit{lastChange} operations
from that point. The algorithm builds on existing breakpoint debugger
technology. We demonstrate the feasibility of the algorithm with
\textit{Querypoint}, an implementation extending the Firebug
JavaScript debugger. \textit{Querypoint} also provides mechanism for
automated bug reproduction, and a novel user interface which
summarizes investigated execution points and collected results.

@@

To start debugging, the developer sets a breakpoint
on line 22 and, once the button is clicked, execution is paused at line
22. Figure ~\ref{fig:example1} shows the Firebug debugger while the
execution is paused. Firebug has several panels (e.g., HTML, CSS,
Script, DOM, etc.) that each demonstrate one aspect of the Web page.
The Script panel contains the list of all loaded source
files and regular debugging facilities such as setting breakpoints and
stepping. To the right of the script panel, the Watch panel shows the program state
where the developer can examine object and variable values. In our case, the
\texttt{myObject.myProperty} value at the paused point is zero. We expected this value to be \texttt{2}.

To apply backward search strategy for locating defects, the developer
first needs to know the origin to the wrong value. To achieve this
goal using breakpoints, the developer should search code to find and
set breakpoint at all possible places that
\texttt{myObject.myProperty} might get a new value.  However, an
object and property can be accessed and changed through different
names and methods. There is no simple way to identify these alias or
even their total number.  The developer can make a good guess and set
breakpoints on lines where the property seems to be changed. Then they
re-execute the program and examine the state looking for values that
may lead to the incorrect value observed at line 22. All this work
must be repeated if a new alias is discovered or if the some
information related to the buggy result was missed while stopped on
one of the breakpoints.

In contrast, we propose a high-level function in debugger,
\textit{lastChange}, which provides the answer without tedious manual
effort from the developer. By right clicking on
\texttt{myObject.myProperty} in the Watch panel, the developer can run
\textit{lastChange} command (Figure ~\ref{fig:example2}). Debugger
re-executes the program and halts again at the breakpoint on line 22.
However, it shows a new QP panel, centered on the source at line 29
(Figure ~\ref{fig:example3}), the point of \textit{lastChange}.  To
the right, the TraceData panels shows values of properties from the
program state when it passed through line 29.  These two panels
resemble the Script and Watch panels, but they show data collected by
the debugger at an execution point which is now past: these are
'traces' or 'logs' of information collected during the re-execution.

Looking at line 29, it seems that something is wrong with
\texttt{myCondition.value} which causes line 29 execution. The
developer examines \texttt{myCondition.value} and it is
\texttt{undefined}. The next step is to know when this property got
this value. To do so, the developer runs \textit{lastChange} command
on \texttt{myCondition.value} at this point. This re-executes the
program and pauses at the reproduction point,  showing line 25-the
place \texttt{oldValue} is assigned to
\texttt{myCondition.value} (Figure ~\ref{fig:example3}). Now it is
clear that the bug occurs because \texttt{oldValue} is
\texttt{undefined} once execution reaches line 25.

As demonstrated in Figure ~\ref{fig:example-points}, the developer
examined three points of execution. The first point, the reproduction point,  was the breakpoint set by the developer.
The second and third points preceded the reproduction point in execution sequence.
All three points-the history
of the search for the defect-are available through the debugger's
interface. On the top of the left panel in Figure ~\ref{fig:example4}
there is an opened list which shows all three examined points. The
first one is the breakpoint on line 22, the second one is the point
which is the last change of \texttt{myObject.myProperty} before
reaching the breakpoint and finally the last one is the point of
execution in which \texttt{myCondition.value} gets the
\texttt{undefined} value. Moreover, the source lines related to these
points are tagged with red-cycle icons.

@@

Notice that in our example, \textit{lastChange} combines some aspects
of breakpoint and of log-based debugging. Like breakpoint debugging,
the developer re-executes a live runtime without changing the source
and without a special execution environment beyond the debugger. The
state of the program memory and the call stack are available at the
lastChange point. Like log-based debugging, the program state and the
call stack are recorded during program execution. We can't halt the
program at \textit{lastChange} because we don't know which is the last
one until we return to the original breakpoint. (In section 5 we
discuss cases were it is possible to pause at lines of \textit{lastChange}).

@@

The \textit{lastChange} algorithm is based on program re-execution of
a program halted on a breakpoint. The breakpoint hit becomes the
\textit{reproduction point}. The algorithm starts when developer
examines the program state at a breakpoint hit and asks for the
\textit{lastChange} of a value.  Debugger sets hooks (a callback
function dependent upon the underlying runtime) on all instructions
that might be the result of \textit{lastChange} query. Then the
debugger re-executes the program and every time a hook hits it
checks for a \textit{change event}. In the case of a change, it stores
part of the program state values.  Once the execution reaches the
reproduction point, it analyzes the collected data and shows the
result.  The program state at the execution point for the last change
event is the \textit{lastChange}.

While \textit{lastChange} itself is not a different kind \textit{breakpoint},
all of the assumptions developers already have for breakpoints hold
for \textit{lastChange}. In particular the re-execution is just the
same operation developers use to debug with breakpoints. Because
\textit{lastChange} dramatically reduces the developers work in
setting and removing breakpoints, the re-execution step will become a
much larger fraction of the debug time. Therefore automatic mechanisms
for re-execution will be corresponding more valuable for debugging
with \textit{lastChange}.

@@

\subsection{\textit{lastChange} on object property}
To simplify the algorithm explanation and defer technical details, we
define two basic operations and later we explain the details of these
two operations. The first operation is \texttt{objectId()}: given a
JavaScript object it returns an integer as its identifier. This
identifier is unique to the object during one execution.  By using an
object id instead of an object reference we allow the garbage
collector to reclaim the space for dead objects just at would in the
absence of the debugger. The second
operation is \texttt{setPropertyChangeHook()}: given a function and a
string, the function is called whenever a property changes and its
name matches the string For example, if the property might be
\texttt{foo} so changes to \texttt{bar.foo} or \texttt{baz.foo} would
call the function.  The callback function receives a reference to the
owner of \texttt{foo}. In the next section, we describe how we use
a slightly different version of this operation which is restricted 
to only created objects in one location instead of all objects.

@@

  function f(){
    var x;
    x++;
    if (!stop) f();
  }

@@

\subsubsection{global and \texttt{with} scopes}
Global scope is the the most outer scope in the scope chain and it is
also referred to as the global object (the \texttt{window}
object in Web pages). This scope is a regular JavaScript object and therefore every
global variable is a property of global object. Similarly,
 \texttt{with} scopes are also regular objects. A \texttt{with}
scope is created by a \texttt{with()} block with an object as the
parameter. Every property of this parameter object is available inside the
block as a variable. \textit{lastChange} treats the case where variable's scope
is global or \texttt{with}, like \textit{lastChange} on an object
property.

\subsubsection{local, closure and \texttt{catch} scopes}
Local scope refers to the most nested scope in the scope
chain and contains the local variables. Closure scope refers to the 
scope which is created for a nested function and contains variables 
defined in the outer block. Catch scope is the scope created in the catch
block of try-catch statements and contains the exception
variable. These scopes are not necessarily regular JavaScript
objects. Scope object is usually a transient copy of a native object. 
Therefore, to track changes to a variable in these scopes we
employ a different approach.

Having the scope chain and the source code, we can map every scope to % perhaps needs more explanation, how?
a code block, enclosed the executing code. In
JavaScript, a code block can be identified by the file url and the 
the block's first instruction program counter. Given this information, debugger is
able to recognize the code block once it is loaded.
Similar to \textit{lastChange} on object property, we define two basic operations: 
\texttt{scopeId()} and \texttt{setVariableChangeHook()}. The first one 
given a scope returns an integer as the scope's identifier. The second one 
given a callback function, a code block and a variable name-
which is defined inside the block-, the function is called
whenever the variable is changed. Figure ~\ref{fig:recursive} demonstrates a case that the same variable
\texttt{x} is changed several times but in different scopes due to the
recursive calls. Without considering the scopes calling \textit{lastChange(x)}
at point \texttt{C} in scope 1 will return point \texttt{B} which is a change of \texttt{x} in 
scope 2 while the right answer is point \texttt{A}. To resolve this issue, an id
is assigned to the scope by \texttt{ScopeId()} operation and the scope id is also stored 
at every variable change event.

@@

\subsection{\textit{lastChange} on \textit{lastChange}}
The algorithm we explained so far needs an enhancement to support
cascades of \textit{lastChange}s. We explain first the issue
and then the solution. Consider the following points:

@@

To address this issue, we use Algorithm \ref{dependency-analysis}, to
create the list of additional data should be collected at a change
event. This algorithm does does a dependency analysis between all
\textit{lastChange} queries and computes the list of additional data
required to be collected at each change event. For example in the
mentioned case, \texttt{baz} id-if it is available- should be
collected at all points \texttt{x} changes. Having this additional
data, \texttt{baz} id at point B, it is enough to look at the list
of \texttt{y} changes and find the last change event with the same 
owner id (Figure ~\ref{fig:lastchange-lastchange}).

@@

An object might be created by one of these statements: object
initializer (\texttt{\{...\}}), \textit{new} operator (\texttt{new 
constructor()}) or function definition statment (\texttt{function()}). 
By parsing the source code we can easily recognize the statements 
create an object. The next step is getting a reference to the new object.

The new object can be assigned to an object property or variable by
an assignment (\texttt{=}). In these cases we keep the assignee statement 
at the left side. The idea is that we create a list of assignee statements 
that the new objects are assigned to them. We set a hook on the creation
line. Once the hook hits, we evaluate all assignee statements. Then 
we do stepping(step-over) and after each step we evaluate the
statements. Every statement which has a new value, we consider the new
value-if it is an object-as a new object. For example in Figure
~\ref{fig:objectCreation}(a), if the creation line number is 20, we
have only one statement which creates a new object and it is assigned
to \texttt{x.y}.

@@

Although this approach is successful in many ordinary cases, there are some 
cases that this approach doesn't work. For example in a case
where the new object is assigned to \texttt{a[++i]} where i is a variable
(Figure ~\ref{fig:objectCreation}(d)), \texttt{a[++i]} is kept in the list
of assignee statements. Obviously, evaluating this statement doesn't return
a reference to the new object. Our prototype implementation does not yet handle these kinds of unusual cases.

@@

\subsection{\texttt{scopeId()} operation}
The debugger sets a hook at the beginning of all code blocks needing
a scope id whenever the block is loaded. The scope id is kept as a variable with name 
\texttt{\_scopeId} in the scope. Whenever the hook hits,  meaning a new scope
is created, \texttt{\_scopeId} is set by calling JavaScript's dynamic compilation 
function \texttt{eval()}. For example, executing \texttt{eval( var \_scopeId = 10)} creates a
variable with name \texttt{\_scopeId} and value \texttt{10} in the
scope of the \texttt{eval()} call, which is our interesting scope. \texttt{scopeId()}
operation returns the value of \texttt{\_scopeId} in the scope.

\subsection{\texttt{setVariableChangeHook()} operation}
Variables defined in local, closure, and catch scopes are only changed in
their scope; that includes the defining scope and scopes nested inside of them.
For example in Figure ~\ref{fig:js-closure}, variable \texttt{var1} in line 11 can only be
changed in lines 10 to 23. Therefore, after locating the function in
which the variable is defined, it is enough to parse the code inside
the function block and set a hook on all lines where the variable is
assigned a new value. 
We also set a hook at the first line of the function which is
corresponding to the line variable is defined. In Figure
~\ref{fig:js-closure}, if the execution is paused at line 17 and the
last change of \texttt{var1} is queried, two hooks on lines 16 and 17
will be set, but if the execution is paused at line 20 , four hooks on
lines 11, 12, 14, 20 will be set. These two cases are different due
the fact that \texttt{var1} in \texttt{firstChild} is a local variable
but in \texttt{secondChild} is a closure variable.

@@

\textit{Querypoint} keeps track of breakpoint hits and single steps. For example, 
if developer query \textit{lastChange} at the third hit of breakpoint \textit{b}, in
re-execution, third hit of breakpoint b recognized as the reproduction point.

To build a test case from one execution, two elements should be carefully 
considered. First, the initial state should be the same as the initial state
of execution. Second, similar actions and events should be applied to
the program during the execution. \textit{Querypoint} supports two mechanism
for automatic re-execution: local-reproduction, record-replay. The first one,
recalls the function at the bottom of call stack with the same parameters. The idea
behind this mechanism is that many bugs in web pages can be reproduced by firing
an event like clicking on a button. The second one, in the first execution, record phase, 
stores the initial page url and captures user actions. In the replay phase, it opens
the same url and applies user actions. The local reproduction mechanism provides
shorter re-execution cycles while the record-replay is more accurate about the
initial state.

>>>>

According to \cite{LaToza}, developers spend about fifty percent of
their time debugging. To fix a bug, developers typically reproduce and
monitor the buggy execution several times to understand the program's
unexpected behavior. Trial and error, guess-work, and analyzing large
and complicated data make debugging difficult and time-consuming. Any
enhancement to debugging techniques and tools can save developers'
time, reduce development costs and improve programs' quality.

A common strategy for locating defects starts from bug symptoms and
works backwards, moving from a point in the program execution where a
value appears to be incorrect back to the point were that value was
set.  Two conventional approaches are used: breakpoint-based and
log-based debugging. Both approaches require tedious steps of
selecting data to be collected, collecting the data, then analyzing
the results.

In breakpoint debugging, developers select data to be collected by
searching through source files and setting breakpoints. To determine
where a value was set incorrectly, a developer must set
breakpoints at all possible points where the value changes. At
every breakpoint, the developer must determine if the location is in
fact related to the questionable value change then study the complex
debugger user interface and memorize values or manually collect
data. As the number of breakpoint hits increases, the process of
checking the program state, collecting data and resuming the execution
becomes cumbersome.

In log-based debugging, developers select data to be collected by
inserting statements for all points of possible change.  While in
breakpoint-based debugging, the whole program state is available to
developer, in log-based debugging, developer has to decide what data
should be collected when inserts the log statement. It is very common
that the developer has to repeat this step several times due to
insufficient collected data, or to wait a long time because too much
data is recorded. Once the adequate data is collected, it still
requires analyzing and understanding. Developers usually end up in
dealing with long log files and analyzing huge collected data.  None
of these approaches assist developer in finding origins to a wrong
value.

To address these problems, we introduce a new functionality in
debuggers, \textit{lastChange}, which locates the origin of a wrong
value. It uses program re-execution similar to the practice
in using breakpoints and log statements.

@@

We refer to the breakpoint used in \textit{lastChange} as the
\textit{reproduction point}: this is a point where the developer knows
reproduces the bug and it is the point were the \textit{lastChange}
algorithm ends. The ability to identify this reproduction point, i.e.
 a test case and a breakpoint after the bug has appeared, is the only
prerequisite for \textit{lastChange} functionality.  This means
\textit{lastChange} can be implemented on top of a conventional
breakpoint debugger and it does not require a special environment to
create identical, instruction by instruction, re-executions.

Our contribution in this paper is the algorithm \textit{lastChange},
which locates the last place a value has changed, gathers other values
from that execution point, and allows \textit{lastChange} operations
from that point. The algorithm builds on existing breakpoint debugger
technology. We demonstrate the feasibility of the algorithm with
\textit{Querypoint}, an implementation extending the Firebug
JavaScript debugger. \textit{Querypoint} also provides mechanism for
automated bug reproduction, and a novel user interface which
summarizes investigated execution points and collected results.

@@

By browsing through the code, the developer can quickly determine that the value displayed on the 
button is set at line 22. To start debugging, the developer sets a breakpoint
on line 22. Once the button is clicked, execution is paused at line
22. Figure ~\ref{fig:example1} shows the Firebug debugger while the
execution is paused. Firebug has several panels (e.g., HTML, CSS,
Script, DOM, etc.) that each demonstrate one aspect of the Web page.
The Script panel contains the list of all loaded source
files and regular debugging facilities such as setting breakpoints and
stepping. To the right of the script panel, the Watch panel shows the program state
where the developer can examine object and variable values. In our case, the
\texttt{myObject.myProperty} value at the paused point is zero. We expected this value to be \texttt{2}.

To apply backward search strategy for locating defects, the developer
first needs to know the origin of the wrong value. To achieve this
goal using breakpoints, the developer should search code to find all possible places that
\texttt{myObject.myProperty} might get a new value and set breakpoint at these locations. However, an
object and property can be accessed and changed through different
names and methods. There is no simple way to identify these alias or
even their total number.  The developer can make a good guess and set
breakpoints on lines where the property seems to be changed. Then they
re-execute the program and examine the state looking for values that
may lead to the incorrect value observed at line 22. All this work
must be repeated if a new alias is discovered or if some
information related to the buggy result was missed while stopped on
one of the breakpoints.

In contrast, we have added a high-level function in debugger,
\textit{lastChange}, which provides the answer without tedious manual
effort from the developer. By right clicking on
\texttt{myObject.myProperty} in the Watch panel, the developer can run
\textit{lastChange} command (Figure ~\ref{fig:example2}). Debugger
re-executes the program and halts again at the breakpoint on line 22.
However, it shows a new panel, called QP, centered on the source at line 29
(Figure ~\ref{fig:example3}), the point of \textit{lastChange}.  To
the right, the TraceData panel shows values of properties of the
program state when it passed through line 29.  These two panels
resemble the Script and Watch panels, but they show data collected by
the debugger at one execution point which is now past: these are
'traces' or 'logs' of information collected during the re-execution.

Looking at line 29, it seems that something is wrong with
\texttt{myCondition.value} which causes line 29 execution. The
developer examines \texttt{myCondition.value} and it is
\texttt{undefined}. The next step is to know when this property got
this value. To do so, the developer reruns \textit{lastChange} command
on \texttt{myCondition.value} at this point. This re-executes the
program and pauses at the reproduction point,  showing line 25-the
place \texttt{oldValue} is assigned to
\texttt{myCondition.value} (Figure ~\ref{fig:example4}). Now it is
clear that the bug occurs because \texttt{oldValue} is
\texttt{undefined} once execution reaches line 25.

As demonstrated in Figure ~\ref{fig:example-points}, the developer has
examined three points of execution. The first point, the reproduction point,  was the breakpoint set by the developer.
The second and third points preceded the reproduction point in execution sequence.
All three points-the history
of the search for the defect-are available through the debugger's
interface. On the top of the left panel in Figure ~\ref{fig:example4}
there is an opened list which shows all three examined points. The
first one is the breakpoint on line 22, the second one is the point
which is when \texttt{myObject.myProperty} changed before
reaching the breakpoint and finally the last one is the point of
execution in which \texttt{myCondition.value} gets the
\texttt{undefined} value. Moreover, the source lines related to these
points are tagged with red-cycle icons.

@@

Notice that in our example, \textit{lastChange} combines some aspects
of breakpoint and of log-based debugging. Like breakpoint debugging,
the developer re-executes a live runtime without changing the source
and without a special execution environment beyond the debugger. The
state of the program memory and the call stack are available at each
lastChange point. Like log-based debugging, the program state and the
call stack are recorded during program execution. We can't halt the
program at \textit{lastChange} because we don't know which point is the last
one until we return to the original breakpoint. (In section 5 we
discuss cases were it is possible to pause at lines of \textit{lastChange}).

@@

The \textit{lastChange} algorithm is based on program re-execution of
a program halted on a breakpoint. The algorithm starts when developer
examines the program state at a breakpoint hit and asks for the
\textit{lastChange} of a value. The breakpoint hit becomes the
\textit{reproduction point}. Debugger sets hooks (a callback
function dependent upon the underlying runtime) on all instructions
that might be the result of \textit{lastChange} query. Then the
debugger re-executes the program and every time a hook hits it
checks for a \textit{change event}. In the case of a change, it stores
part of the program state values.  Once the execution reaches the
reproduction point, it analyzes the collected data and shows the
result.  The program state at the execution point of the last change
event is the \textit{lastChange}.

As \textit{lastChange} itself is not a different kind of \textit{breakpoint},
all of the assumptions developers already have for breakpoints hold
for \textit{lastChange}. In particular the re-execution is just the
same operation developers use to debug with breakpoints. Because
\textit{lastChange} dramatically reduces the developers work in
setting and removing breakpoints, the re-execution step will become a
much larger fraction of the debug time. Therefore automatic mechanisms
for re-execution will be corresponding more valuable for debugging
with \textit{lastChange}.

@@

\subsection{\textit{lastChange} on object property}
To simplify the algorithm explanation and defer technical details, we
define two basic operations and later we explain the details of these
two operations. The first operation is \texttt{objectId()}: given a
JavaScript object it returns an integer as its identifier. This
identifier is unique to the object during one execution.  By using an
object id instead of an object reference we allow the garbage
collector to reclaim the space for dead objects just as it would in the
absence of the debugger. The second
operation is \texttt{setPropertyChangeHook()}: given a function and a
string, the function is called whenever a property changes and its
name matches the string For example, if the is 
\texttt{foo} so changes to \texttt{bar.foo} or \texttt{baz.foo} would
call the function.  The callback function receives a reference to the
owner of \texttt{foo}. In the next section, we describe how we use
a slightly different version of this operation which is restricted 
to the created objects in one location instead of all objects.

@@

  function f(){
    var x = 0;
    x++;
    if (!stop) f();
  }

@@

\subsubsection{global and \texttt{with} scopes}
Global scope is the the most outer scope in the scope chain and it is
also referred to as the global object (the \texttt{window}
object in Web pages). This scope is a regular JavaScript object and therefore every
global variable is a property of global object. Similarly,
 \texttt{with} scopes are also regular objects. A \texttt{with}
scope is created by a \texttt{with()} block with an object as the
parameter. Every property of this parameter object is available inside the
block as a variable. \textit{lastChange} treats the case where variable's scope
is global or \texttt{with}, like it does on an object
property.

\subsubsection{local, closure and \texttt{catch} scopes}
Local scope refers to the most nested scope in the scope
chain which contains the local variables. Closure scope refers to the 
scope which is created for a nested function and contains variables 
defined in the outer block. Catch scope is the scope created in the catch
block of try-catch statements and contains the exception
variable. These scopes are not necessarily regular JavaScript
objects. Scope object is usually a transient copy of a native object. 
Therefore, to track changes to a variable in these scopes we
employ a different approach.

Having the scope chain and the source code, we can map every scope to % perhaps needs more explanation, how?
a code block, enclosed the executing code. In
JavaScript, a code block can be identified by the file url and  
the block's first instruction program counter. Given this information, the debugger is
able to recognize the code block once it is loaded.
Similar to \textit{lastChange} on object property, we define two basic operations: 
\texttt{scopeId()} and \texttt{setVariableChangeHook()}. The first one, 
given a scope returns an integer as the scope's identifier. The second one, 
given a callback function, a code block and a variable name-
which is defined inside the block-, the function is called
whenever the variable is changed. Figure ~\ref{fig:recursive} demonstrates a case in which the same syntactically variable
\texttt{x} is changed several times but in different scopes due to the
recursive calls. Without considering the scopes calling \textit{lastChange(x)}
at point \texttt{C} in scope 1 will return point \texttt{B} which is a change of \texttt{x} in 
scope 2 while the right answer is point \texttt{A}. To resolve this issue, an id
is assigned to the scope by \texttt{ScopeId()} operation and the scope id is also stored 
at every variable change event.

@@

\subsection{\textit{lastChange} on \textit{lastChange}}
The algorithm we explained so far needs an enhancement to support
cascades of \textit{lastChange}s. We first explain the issue
and then the solution. Consider the following points:

@@

To address this issue, we use Algorithm \ref{dependency-analysis}, to
create the list of additional data (object id or scope id) should be collected at a change
event. This algorithm does a dependency analysis between all
\textit{lastChange} queries and computes the list of additional data
required to be collected at each change event. For example in the
mentioned case, \texttt{baz} id-if it is available- should be
collected at all points \texttt{x} changes. Having this additional
data, \texttt{baz} id at point B, it is enough to look at the list
of \texttt{y} changes and find the last change event with the same 
owner id (Figure ~\ref{fig:lastchange-lastchange}).

@@

An object might be created by one of these statements: object
initializer (\texttt{\{...\}}), \textit{new} operator (\texttt{new 
constructor()}) or function definition statment (\texttt{function()}). 
By parsing the source code we can recognize the statements that
create an object. The next step is getting a reference to the new object.

The new object can be assigned to an object property or a variable by
an assignment (\texttt{=}). In these cases we keep the assignee statement 
at the left side. The idea is that we create a list of assignee statements 
that the new objects are assigned to. We set a hook on the creation
line. Once the hook hits, we evaluate all assignee statements. Then 
we do stepping(step-over) and after each step we evaluate the
statements. Every statement which has a new value, we consider the new
value-if it is an object-as a new object. For example in Figure
~\ref{fig:objectCreation}(a), if the creation line number is 20, we
have only one statement which creates a new object and it is assigned
to \texttt{x.y}.

@@

Although this approach is successful in many ordinary cases, there are some 
cases where this approach doesn't work. For example in a case
where the new object is assigned to \texttt{a[++i]} where i is a variable
(Figure ~\ref{fig:objectCreation}(d)), \texttt{a[++i]} is kept in the list
of assignee statements. Obviously, evaluating this statement doesn't return
a reference to the new object. Our prototype implementation does not handle these kinds of unusual cases yet.

@@

\subsection{\texttt{scopeId()} operation}
The debugger sets a hook at the beginning of all code blocks needing
a scope id. The scope id is kept as a variable with name 
\texttt{\_scopeId} in the scope. Whenever the hook is hit,  meaning a new scope
is created, \texttt{\_scopeId} is set by calling JavaScript's dynamic compilation 
function \texttt{eval()}. For example, executing \texttt{eval( var \_scopeId = 10)} creates a
variable with name \texttt{\_scopeId} and value \texttt{10} in the
scope of the \texttt{eval()} call, which is our interesting scope. \texttt{scopeId()}
operation returns the value of \texttt{\_scopeId} in the scope.

\subsection{\texttt{setVariableChangeHook()} operation}
Variables defined in local, closure, and catch scopes are only changed in
their scope; that includes the defining scope and scopes nested whithin them.
For example in Figure ~\ref{fig:js-closure}, variable \texttt{var1} in line 11 can only be
changed in lines 10 to 23. Therefore, after locating the function in
which the variable is defined, it is enough to parse the code inside
the function block and set a hook on all lines where the variable is
assigned a new value. 
We also set a hook at the first line of the function which is
corresponding to the line where the variable is defined. In Figure
~\ref{fig:js-closure}, if the execution is paused at line 17 and the
last change of \texttt{var1} is queried, two hooks on lines 16 and 17
will be set, but if the execution is paused at line 20 , four hooks on
lines 11, 12, 14, 20 will be set. These two cases are different due
the fact that \texttt{var1} in \texttt{firstChild} is a local variable
but in \texttt{secondChild} is a closure variable.

@@

\textit{Querypoint} keeps track of breakpoint hits and single steps. For example, 
if the developer queries \textit{lastChange} at the third hit of breakpoint \textit{b}, in
re-execution, the third hit is recognized as the reproduction point.

To build a test case from one execution, two elements should be carefully 
considered. First, the initial state should be the same as the initial state
of execution. Second, similar actions and events should be applied to
the program during the execution. \textit{Querypoint} supports two mechanism
for automatic re-execution: local-reproduction, record-replay. The first one,
recalls the function at the bottom of the call stack with the same parameters. The idea
behind this mechanism is that many bugs in web pages can be reproduced by firing
an event like clicking on a button. The second one, in the first execution, record phase, 
stores the initial page url and captures user actions. In the replay phase, it opens
the same url and applies user actions. The local reproduction mechanism provides
shorter re-execution cycles while the record-replay is more accurate about the
initial state.
